---
title: "add_vax"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(SMILE)
library(tidyverse)
library(cowplot)
```

We incorporate vaccination into the model as a term that modifies the probability of moving from compartment I to D or M. This means, vaccination is incorporated into the model as having an effect over survival probability, by changing the probability of moving from the infected compartment to the LIZ (death) or M (Immune) compartments.

As a simple exercise, we can consider a population over the course of 10 years, where the vaccination rate is held constant. In the case of a constant vaccination rate, we are essentially just changing the survival probability, and it remains the same over the course of the simulation.

First, we set our parameters:
```{r}
# base parameters
my_years <- 10
my_tau <- 1
my_theta <- 100

# seasonal forcing parameters
my_b_fixed <- 0.001
my_b0 <- -30
my_b1 <- 0.85
my_outbreak_period <- 3*52

# vaccination parameters
my_zeta_no_vax <- 0.88 # survival no vaccination
my_fixed_vax <- rep(0.5, my_years) # fixed vaccination
my_yearly_vax <- runif(my_years, min = 0, max = 0.5) # random vaccination rates by year
my_beta0 <- -5
my_beta1 <- 10

```


```{r}
smile_main(b0 = my_b0, b1 = my_b1, period = my_outbreak_period,
           theta = my_theta, tau = my_tau, years = my_years,
           vax = my_fixed_vax, beta_0 = my_beta0, beta_1 = my_beta1, output_df = TRUE) -> fixed_vax_sim

head(fixed_vax_sim)
```

We create a simple function to help us visualize the data we just generated 

```{r}
build_SMILE_plots <- function(smile_output_as_df, facet_rows = 2) {
  smile_output_as_df %>% 
    pivot_longer(-week, names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "Z"))) %>% 
    ggplot(aes(x = week, y = n)) +
    facet_wrap(~compartment, scales = "free", nrow = facet_rows) +
    geom_path() +
    theme_bw()
  
}
```

```{r}
fixed_vax_sim |> build_SMILE_plots()
```


# We can compare this to variable vaccination rates:


```{r}
smile_main(b0 = my_b0, b1 = my_b1, period = my_outbreak_period,
           theta = my_theta, tau = my_tau, years = my_years,
           vax = my_yearly_vax, beta_0 = my_beta0, beta_1 = my_beta1, output_df = TRUE) -> var_vax_sim

head(var_vax_sim)
```
```{r}
var_vax_sim |> build_SMILE_plots()
```


From this simulation we observe that with drops in vaccination, we have a direct increase in the number of deaths and spores in the environment.


