---
title: "main_function"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Intro 

Start by loading our package. In this article we want to show how the updated function works to recreate the same simulation scenarios tested in the first iteration of this code. Previously, multiple functions were needed to incorporate population dynamics or seasonality, or both. In this example, we will show you how to set the parameters to the package's main function `SMILE::smile_main()` and generate the same scenarios as before. 

```{r setup}
set.seed(978645)
library(SMILE)
```

The following packages are only needed for visualization and data wrangling purposes:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
```

The original code is found in the [first commit for this repository](https://github.com/UFSEERLab/SMILE/tree/e176302c90e16205b0211906e1ef93e641776407) with the defined functions to evaluate the model under different assumptions of population dynamics or seasonal forcing. This is code show in also shown before each of the scenarios. 

```{r OG_fxs}
# OG functions 

# Author: Juan Pablo Gomez
# Version 1.1.
# Last revised: March 7 2018.

#################################
# Infection probability based on Ponciano and Capistran 2011.
lambda.t	<-	function(theta,tau,b,E){
	
	1-(theta/(theta+b*E))^tau
	
}

# Introducing seasonality in the infection probability through b.

b.season	<-	function(b0,b1,period,t){
	
	exp(b0*(1+b1*cos((2*pi*t)/period)))
	
}

# Density dependent reproduction

rho.n	<-	function(N){
	
	0.41/(1+(N/5000)^(10))
	
}

# First case of R0

r0	<-	function(b,E,theta,tau){
	
	(b*E*tau)/theta
	
}

# Function to simulate climatic variable
# a = upper asymptote
# c = lower asymptote
# b = time required to reach half of the climatic maxima
# d = time at which the mid point between max and minimum climate is reached.
clim.func	<-	function(a,c,d,d2,b,b2,t,sd){
	
	if(missing(b2)){b2 <- b}
	if(missing(c)){c<-0}
	
	det.clim	<-	 ((a-c)/(1+exp((d-t)/b)) - (a-c)/(1+exp((d2-t)/b2))) + c
	rand.clim	<-	rnorm(length(t),det.clim,sd=sd)
	return(rand.clim)
	
}

# Function used to estimate seasonality in infection probability using climatic variables
# wt.bar = mean of climatic covariable
# b0 = as in b.season
# b1 = as in b.season
# kw = scaling coefficient
# t = time
# period = as in b.season

infect2clim	<-	function(wt.bar,b0,b1,kw,t,period){
	
	wt.bar + ((b0*b1)/kw)*cos((2*pi*t)/period)
	
}

```

```{r echo=FALSE}
# function to make the plots comparing original function to new function:

build_comp_plots <- function(og_sims, new_sims) {
  
  sim5.0 <- og_sims
  sim5_1 <- new_sims
  
      # Susceptibles
    sim5.0$Suceptibles %>% tibble() %>% rownames_to_column() %>% purrr::set_names(c("week", "S")) %>% mutate(fx = "OG") %>% 
      mutate(week = as.numeric(week)) -> S_comp
      
    S_comp %>% 
      bind_rows(sim5_1$Suceptibles %>% tibble() %>% rownames_to_column() %>%
                  purrr::set_names(c("week", "S")) %>% mutate(fx = "NEW", week = as.numeric(week))) %>% 
      ggplot(aes(x = week, y = S)) +
      facet_wrap(~fx) +
      geom_line() +
      theme_bw() -> S_plot
    
    # Immune
    sim5.0$Immune %>% tibble() %>% rownames_to_column() %>% purrr::set_names(c("week", "M")) %>% mutate(fx = "OG") %>% 
      mutate(week = as.numeric(week)) -> M_comp
      
    M_comp %>% 
      bind_rows(sim5_1$Immune %>% tibble() %>% rownames_to_column() %>%
                  purrr::set_names(c("week", "M")) %>% mutate(fx = "NEW", week = as.numeric(week))) %>% 
      ggplot(aes(x = week, y = M)) +
      facet_wrap(~fx) +
      geom_line() +
      theme_bw() -> M_plot
    
    # Infected
    sim5.0$Infected %>% tibble() %>% rownames_to_column() %>% purrr::set_names(c("week", "I")) %>% mutate(fx = "OG") %>% 
      mutate(week = as.numeric(week)) -> I_comp
      
    I_comp %>% 
      bind_rows(sim5_1$Infected %>% tibble() %>% rownames_to_column() %>%
                  purrr::set_names(c("week", "I")) %>% mutate(fx = "NEW", week = as.numeric(week))) %>% 
      ggplot(aes(x = week, y = I)) +
      facet_wrap(~fx) +
      geom_line() +
      theme_bw() -> I_plot
    
    # LIZ
    sim5.0$LIZ %>% tibble() %>% rownames_to_column() %>% purrr::set_names(c("week", "L")) %>% mutate(fx = "OG") %>% 
      mutate(week = as.numeric(week)) -> L_comp
      
    L_comp %>% 
      bind_rows(sim5_1$LIZ %>% tibble() %>% rownames_to_column() %>%
                  purrr::set_names(c("week", "L")) %>% mutate(fx = "NEW", week = as.numeric(week))) %>% 
      ggplot(aes(x = week, y = L)) +
      facet_wrap(~fx) +
      geom_line() +
      theme_bw() -> L_plot
    
    
    # E
    sim5.0$E %>% tibble() %>% rownames_to_column() %>% purrr::set_names(c("week", "E")) %>% mutate(fx = "OG") %>% 
      mutate(week = as.numeric(week)) -> E_comp
      
    E_comp %>% 
      bind_rows(sim5_1$E %>% tibble() %>% rownames_to_column() %>%
                  purrr::set_names(c("week", "E")) %>% mutate(fx = "NEW", week = as.numeric(week))) %>% 
      ggplot(aes(x = week, y = E)) +
      facet_wrap(~fx) +
      geom_line() +
      theme_bw() -> E_plot
    
    plot_grid(S_plot, M_plot, I_plot, L_plot, E_plot, ncol = 3)

}
```

# smile1
No age structuring, infection probability is not seasonal 

```{r}
# Function of SMILE without any age structuring
# This is the most basic SMILE function in which I have assumed no population dynamics
# and no deaths other than disease related. 
# Fixed parameters are: 
# alpha= 1/52 Probability of an immune individual to become Suceptible 
# zeta = 0.88 Probability that an Infected individual becomes immune
# gamma= 0.9868 Spore decay probability
# psi = 1 Number of spores in one LIZ

# Variable parameters
# b: Number of infections caused by one LIZ if there was no dispersion effort
# tau and theta: Shape and Rate parameters of the gamma distribution defining dispersion effort.

smile1	<-	function(b,theta,tau,years){
			
	# Fixed parameters
	alpha	<-	1/52
	zeta	<-	0.88
	gamma	<-	0.9868
	psi		<-	1	 
	
	n.weeks		<-	years*52 + 1
	S<-	M<-	I<-	L<-E <-N<-lambda<- array(0,dim=c(n.weeks),dimnames=list(1:(n.weeks)))
	N[1]		<-	5000
	S[1]		<-	N[1]
	
	L[1]	<-	1; E[1]	<-	L[1]*psi	
	
	for(t in 2:n.weeks){
		
		tm1	<-	t-1
		lambda[tm1]	<-	lambda.t(theta=theta,tau=tau,b=b,E=E[tm1])
		
		S[t]	<-	(S[tm1]*(1-lambda[tm1])) + M[tm1]*1/52
		I[t]	<-	S[tm1]*lambda[tm1]
		M[t]	<-	I[tm1]*zeta + M[tm1]*(1-(1/52))
		L[t]	<-	I[tm1]*(1-zeta)
		E[t]	<-	psi*L[tm1] + E[tm1]*gamma
		N[t]	<-	S[t]+M[t]
			
	}
	

	results		<-	list(Suceptibles=S[-1]
						,Immune=M[-1]
						,Infected=I[-1]
						,LIZ=L[-1]
						,Environment=E[-1]
						,lambda=lambda[-1])

	return(results)
}
```

```{r}
tau=10
b=0.001

# Case 1.0: Low Dispersion. Mean dispersion effort is 0.1
theta=100
sim1.0	<-	smile1(b=b,tau=tau,theta=theta,years=10)
```

To recreate this scenario with the new function, we use the following parameters. Note the argument `age_struc` is set to `FALSE`.

```{r}
sim1_1 <- smile_main(tau = tau, b_fixed = b, theta = theta, years = 10, age_struc = FALSE)
# sim1_1 <- smile_main(tau = 10, b_fixed = 0.001, theta = 100, years = 10, age_struc = FALSE)
```


```{r echo=FALSE, fig.width=12, fig.height=4}
build_comp_plots(sim1.0, sim1_1)
```

# smile2
This function simulates anthrax disease dynamics without host population dynamics but infection probability has seasonal forcing.

```{r}
# This function simulates anthrax disease dynamics without host population dynamics
# but infection probability has seasonal forcing.
# Fixed parameter are the same as in smile1 function. 
# Variable parameters tau and theta are also the same and b0, b1 and period 
# introduce seasonality to the infection probability by modifying b. In this function
# I assume that seasonality is product of an exponential cosine function where the intensity 
# of the seasonality is given by b0*b1 and period gives the periodicity of the outbreak
# tau and theta: Shape and Rate parameters of the gamma distribution defining dispersion effort.
# period: is the period in weeks of the exponential sinusoid function

smile2	<-	function(b0,b1,period,theta,tau,years){
			
	# Fixed parameters
	alpha	<-	1/52
	zeta	<-	0.88
	gamma	<-	0.9868
	psi		<-	1	 
	
	n.weeks		<-	years*52 + 1
	S<-	M<-	I<-	L<-E <-N<-lambda<- array(0,dim=c(n.weeks),dimnames=list(1:(n.weeks)))
	N[1]		<-	5000
	S[1]		<-	N[1]
	
	L[1]	<-	1; E[1]	<-	L[1]*psi	
	
	for(t in 2:n.weeks){
		
		tm1	<-	t-1
		b	<-	b.season(b0=b0,b1=b1,period=period,t=t)
		lambda[tm1]	<-	lambda.t(theta=theta,tau=tau,b=b,E=E[tm1])
		
		S[t]	<-	(S[tm1]*(1-lambda[tm1])) + M[tm1]*1/52
		I[t]	<-	S[tm1]*lambda[tm1]
		M[t]	<-	I[tm1]*zeta + M[tm1]*(1-(1/52))
		L[t]	<-	I[tm1]*(1-zeta)
		E[t]	<-	psi*L[tm1] + E[tm1]*gamma
		N[t]	<-	S[t]+M[t]
			
	}
	

	results		<-	list(Suceptibles=S[-1]
						,Immune=M[-1]
						,Infected=I[-1]
						,LIZ=L[-1]
						,Environment=E[-1]
						,lambda=lambda[-1])

	return(results)
}
```

```{r}
# Case 2: No age structuring, infection probability has seasonal forcing
tau=10
b0=-30
b1=0.85
period=3*52

# Case 2.0: Low dispersion
theta=100
sim2.0	<-	smile2(b0=b0,b1=b1,period=period,tau=tau,theta=theta,years=10)
```

```{r}

sim2_1 <- smile_main(b0 = b0, b1 = b1, period = period, theta = theta, tau = tau, years = 10, age_struc = FALSE)
# sim2_1 <- smile_main(b0 = -30, b1 = 0.85, period = 3*52, theta = 100, tau = 10, years = 10, age_struc = FALSE)
```


```{r echo=FALSE, fig.width=12, fig.height=4}
build_comp_plots(sim2.0, sim2_1)
```

# smile3
Only births with no seasonal infection probability 

```{r}
# In this function we introduce population dynamics only as births at the begining of the year
# determined by density dependent reproduction with a carrying capacity of 5000.

smile3	<-	function(b,theta,tau,years){
			
	# Fixed parameters
	alpha	<-	1/52
	zeta	<-	0.88
	gamma	<-	0.9868
	psi		<-	1	 
	
	n.weeks		<-	years*52 + 1
	S<-	M<-	I<-	L<-E <-N<-lambda<- array(0,dim=c(n.weeks),dimnames=list(1:(n.weeks)))
	N[1]		<-	5000
	S[1]		<-	N[1]	
	L[1]	<-	1; E[1]	<-	L[1]*psi	
	
	for(t in 2:n.weeks){
				
		tm1	<-	t-1
		lambda[tm1]	<-	lambda.t(theta=theta,tau=tau,b=b,E=E[tm1])
		births.happen	<-	as.numeric(t%%52==0)
		rep.prob	<-	rho.n(N[tm1])
		
		S[t]	<-	(S[tm1]*(1-lambda[tm1])) + M[tm1]*1/52 + rep.prob*(N[tm1])*births.happen
		I[t]	<-	S[tm1]*lambda[tm1]
		M[t]	<-	I[tm1]*zeta + M[tm1]*(1-(1/52))
		L[t]	<-	I[tm1]*(1-zeta)
		E[t]	<-	psi*L[tm1] + E[tm1]*gamma
		N[t]	<-	S[t]+M[t]
			
	}
	

	results		<-	list(Suceptibles=S[-1]
						,Immune=M[-1]
						,Infected=I[-1]
						,LIZ=L[-1]
						,Environment=E[-1]
						,lambda=lambda[-1])

	return(results)
}
```

```{r}
# Case 3: With births and no seasonal infection probability
tau=10
b=0.001

# Case 3.0: Low Dispersion. Mean dispersion effort is 0.1
theta=100
sim3.0	<-	smile3(b=b,tau=tau,theta=theta,years=10)
```

If we don't consider seasonal forcing, then b0 and b1 are left as null (the default) and we give the parameter b_fixed for fixed infection probability. When the argument `age_struc` is not set to `FALSE`, the function will use the default parameters of `K` and `rho_pop` to estimate birth pulses every year. These default parameters of carrying capacity and reproductive rate can be changed easily by adding them to the function as arguments. 
```{r}
sim3_1 <- smile_main(tau = tau, b_fixed = b, theta = theta, years = 10)
# sim3_1 <- smile_main(tau = 10, b_fixed = 0.001, theta = 100, years = 10)
```


```{r echo=FALSE, fig.width=12, fig.height=4}
build_comp_plots(sim3.0, sim3_1)
```


# smile4 
 With births only and seasonal forcing of infection 
 
```{r}
# smile 4 function introduces population dynamics as in smile3 function but uses 
# seasonal forcing for infection probability.

smile4	<-	function(b0,b1,period,theta,tau,years){
			
	# Fixed parameters
	alpha	<-	1/52
	zeta	<-	0.88
	gamma	<-	0.9868
	psi		<-	1	 
	
	n.weeks		<-	years*52 + 1
	S<-	M<-	I<-	L<-E <-N<-lambda<- array(0,dim=c(n.weeks),dimnames=list(1:(n.weeks)))
	N[1]		<-	5000
	S[1]		<-	N[1]	
	L[1]	<-	1; E[1]	<-	L[1]*psi	
	
	for(t in 2:n.weeks){
				
		tm1	<-	t-1
		b	<-	b.season(b0,b1,period,t)
		lambda[tm1]	<-	lambda.t(theta=theta,tau=tau,b=b,E=E[tm1])
		births.happen	<-	as.numeric(t%%52==0)
		rep.prob	<-	rho.n(N[tm1])
		
		S[t]	<-	(S[tm1]*(1-lambda[tm1])) + M[tm1]*1/52 + rep.prob*(N[tm1])*births.happen
		I[t]	<-	S[tm1]*lambda[tm1]
		M[t]	<-	I[tm1]*zeta + M[tm1]*(1-(1/52))
		L[t]	<-	I[tm1]*(1-zeta)
		E[t]	<-	psi*L[tm1] + E[tm1]*gamma
		N[t]	<-	S[t]+M[t]
			
	}
	

	results		<-	list(Suceptibles=S[-1]
						,Immune=M[-1]
						,Infected=I[-1]
						,LIZ=L[-1]
						,Environment=E[-1]
						,lambda=lambda[-1])

	return(results)
}

```
 
```{r}
tau=10
b0=-30
b1=0.85
period=3*52

# Case 4.0: Low Dispersion. Mean dispersion effort is 0.1
theta=100
sim4.0	<-	smile4(b0=b0,b1=b1,period=period,tau=tau,theta=theta,years=10)
```

In the function we change the parameter for deaths from non-disease causes to `sigmaa = 1`, since the smile4 function does not consider deaths from other causes. It is not necessary to change this parameter though, since the default in the function is already set to 1. 
```{r}
sim4_1 <- smile_main(b0 = b0, b1 = b1, period = period, theta = theta, tau = tau, years = 10, sigmaa = 1)
# sim4_1 <- smile_main(b0 = -30, b1 = 0.85, period = 3*52, theta = 100, tau = 10, years = 10, sigmaa = 1)
```

```{r echo=FALSE, fig.width=12, fig.height=4}
build_comp_plots(sim4.0, sim4_1)
```

 

# smile5
This one incorporates both host population dynamics and seasonal forcing, but additionally puts in the death by other causes.

```{r}
# smile5 is the most realistic function of the group since it incorporates both births and deaths from other causes
# than the disease in population dynamics. Deaths are given by a 1- sigmaa = 1 - 0.92^(1/52)
smile5	<-	function(b0,b1,period,theta,tau,years){
			
	# Fixed parameters
	alpha	<-	1/52
	zeta	<-	0.88
	gamma	<-	0.9868
	sigmaa	<-	0.92^(1/52)
	psi		<-	1	 
	
	n.weeks		<-	years*52 + 1
	S<-	M<-	I<-	L<-E <-N<-lambda<- array(0,dim=c(n.weeks),dimnames=list(1:(n.weeks)))
	N[1]		<-	5000
	S[1]		<-	N[1]	
	L[1]	<-	1; E[1]	<-	L[1]*psi	
	
	for(t in 2:n.weeks){
				
		tm1	<-	t-1
		b	<-	b.season(b0,b1,period,t)
		lambda[tm1]	<-	lambda.t(theta=theta,tau=tau,b=b,E=E[tm1])
		births.happen	<-	as.numeric(t%%52==0)
		rep.prob	<-	rho.n(N[tm1])
		
		S[t]	<-	(S[tm1]*(1-lambda[tm1]))*sigmaa + M[tm1]*sigmaa*1/52 + rep.prob*(N[tm1])*births.happen
		I[t]	<-	S[tm1]*lambda[tm1]
		M[t]	<-	I[tm1]*zeta + M[tm1]*sigmaa*(1-(1/52))
		L[t]	<-	I[tm1]*(1-zeta)
		E[t]	<-	psi*L[tm1] + E[tm1]*gamma
		N[t]	<-	S[t]+M[t]
			
	}
	

	results		<-	list(Suceptibles=S[-1]
						,Immune=M[-1]
						,Infected=I[-1]
						,LIZ=L[-1]
						,Environment=E[-1])

	return(results)
}

```


```{r}
tau=10
b0=-30
b1=0.85
period=3*52

# Case 5.0: Low Dispersion. Mean dispersion effort is 0.1
theta=10
sim5.0		<-	smile5(b0=b0,b1=b1,period=period,tau=tau,theta=theta,years=10)
```

```{r}
sim5_1 <- smile_main(b0 = b0, b1 = b1, period = period, theta = theta, tau = tau, years = 10, sigmaa = 0.92^(1/52))
# sim5_1 <- smile_main(b0 = -30, b1 = 0.85, period = 3*52, theta = 10, tau = 10, years = 10, sigmaa = 0.92^(1/52))
```


```{r echo=FALSE, fig.width=12, fig.height=4}
build_comp_plots(sim5.0, sim5_1)
```

# Stochastic version of the last one:

```{r}
# Final function is the stochastic version of smile5 function. It incorporates births, deaths by disease
# and other causes and seasonal forcing of the infection probability. The stochasticity is introduced by 
# assuming that Infected, Immune and LIZ are all binomial random variables. I allowed stochasticity in the 
# number of spores introduced to the environment by assuming that it is poisson random variable with mean
# given by the number of diseas deaths times a constant. Virulence of spores that remain virulent in the 
# environment is also a binomial random variable with with probability of succes gamma = 0.9868.

smile5.stoch	<-	function(b0,b1,period,theta,tau,years){
			
	# Fixed parameters
	alpha	<-	1/52
	zeta	<-	0.88
	gamma	<-	0.9868
	sigmaa	<-	0.92^(1/52)
	psi		<-	1	 
	
	n.weeks		<-	years*52 + 1
	S<-	M<-	I<-	L<-E <-N<-lambda<- array(0,dim=c(n.weeks),dimnames=list(1:(n.weeks)))
	N[1]		<-	5000
	S[1]		<-	N[1]	
	L[1]	<-	1; E[1]	<-	rpois(1,L[1]*psi	)
	
	for(t in 2:n.weeks){
				
		tm1	<-	t-1
		b	<-	b.season(b0,b1,period,t)
		lambda[tm1]	<-	lambda.t(theta=theta,tau=tau,b=b,E=E[tm1])
		births.happen	<-	as.numeric(t%%52==0)
		rep.prob	<-	rho.n(N[tm1])
		
		I[t]	<-	rbinom(1,S[tm1],lambda[tm1])
		births	<-	rbinom(1,N[tm1],rep.prob)*births.happen
		M.surv	<-	rbinom(1,M[tm1],sigmaa)
		M.recov	<-	rbinom(1,M.surv,alpha)
		M.new	<-	rbinom(1,I[tm1],zeta)
		S[t]	<-	rbinom(1,S[tm1]-I[t],sigmaa) + M.recov + births
		M[t]	<-	M.new + (M.surv - M.recov)
		L[t]	<-	(I[tm1]-M.new)
		E[t]	<-	rpois(1,psi*L[tm1]) + rbinom(1,E[tm1],gamma)
		N[t]	<-	S[t]+M[t]
			
	}
	

	results		<-	list(Suceptibles=S[-1]
						,Immune=M[-1]
						,Infected=I[-1]
						,LIZ=L[-1]
						,Environment=E[-1])

	return(results)
}

```


```{r}
stochastic.sims	<-	list()
LIZ.stoch		<-	matrix(0,nrow=100,ncol=520)
for(i in 1:100){
  
	stochastic.sims[[i]]	<-	smile5.stoch(b0=b0,b1=b1,period=period,tau=tau,theta=theta,years=10)
	LIZ.stoch[i,]			<-	stochastic.sims[[i]]$LIZ
	
}
```

```{r}

stochastic_sims	<-	list()
LIZ_stoch		<-	matrix(0,nrow=100,ncol=520)
for(i in 1:100){
  
	stochastic_sims[[i]]	<-	smile_main(b0 = b0, b1 = b1, period = period, theta = theta, tau = tau, years = 10, sigmaa = 0.92^(1/52), 
	                                   stochastic = TRUE)
	LIZ_stoch[i,]			<-	stochastic_sims[[i]]$LIZ
	
}

```


```{r}
as.data.frame(LIZ.stoch) %>% rownames_to_column(var = "sim") %>% 
  pivot_longer(-sim, names_to = "week", values_to = "n_LIZ") %>% 
  mutate(week = str_remove(week, "V"), week = as.numeric(week)) %>% 
  mutate(model = "OG") -> og_stochastic

as.data.frame(LIZ_stoch) %>% rownames_to_column(var = "sim") %>% 
  pivot_longer(-sim, names_to = "week", values_to = "n_LIZ") %>% 
  mutate(week = str_remove(week, "V"), week = as.numeric(week)) %>% 
  mutate(model = "NEW") -> new_stochastic
```

Very similar results, considering this is the stochastic versions. It makes sense. 
```{r}
bind_rows(og_stochastic, new_stochastic) %>% 
  # filter(sim %in% 10:15) %>% 
  ggplot(aes(x = week, y = n_LIZ, group = sim)) +
  facet_wrap(~model) +
  # facet_grid(sim~model) +
  geom_line() +
  theme_minimal()
```

# Estimation 
Estimation of parameters using the last scenario and stochastic simulations

```{r}
# Functions for estimating the parameters of the SMILE model b0, b1, period, theta and tau.
# The estimation assumes that the observed LIZ are poisson distributed product of observation error
# and not process error as properly described in the smile5.stoch function. This is a set of two functions
# the first one gives the negative loglikelihood of the model that will be minimized using the optim function.
# The second one is a wrapper of the optim function that returns the maximum likelihood estimates of the parameters.
# Period must be in weeks and SMILE.obs must follow exactly the formating of the output of one of the above SMILE 
# functions.

LIZ.negll	<-	function(pars=c(theta,tau,b0,b1),period,years,SMILE.obs){
		
	theta	<-	exp(pars[1])
	tau		<-	exp(pars[2])
	b0		<-	pars[3]
	b1		<-	pars[4]
	# Likelihood of infection seasonality as a function of climate
	
	#wt.bar.hat	<-	mean(clim)
	#clim.pred	<-	infect2clim(wt.bar=wt.bar.hat,b0=b0,b1=b1,kw=kw,t=1:length(clim),period=period)
	#clim.ssq	<-	sum((clim-clim.pred)^2)
	
	SMILE.pred	<-	smile5(b0,b1,period,theta,tau,years)
	ts2keep		<-	names(SMILE.pred)%in%names(SMILE.obs)
	SMILE.pred	<-	SMILE.pred[which(ts2keep==TRUE)]
	loglik.ls	<-	vector("list",length(SMILE.obs))
	
	for(i in 1:length(SMILE.obs)){
		
		loglik.ls[[i]]	<-	dpois(SMILE.obs[[i]],SMILE.pred[[i]],log=TRUE)
		
	}
	
	loglik.ls	<-	lapply(loglik.ls,sum)
	
	loglik	<-	sum(unlist(loglik.ls))
	
	if(!is.finite(loglik)){loglik=-.Machine$double.xmax}
	negll	<-	-loglik #+ clim.ssq
	return(negll)
	
}

SMILE.param.estim	<-	function(b0,b1,theta,tau,period,SMILE.obs,method="BFGS"){
	
	years		<- 	length(SMILE.obs[[1]])/52
	pars		<-	c(theta,tau,b0,b1)
			
	optim.res	<-	optim(par=pars,fn=LIZ.negll,method=method,SMILE.obs=SMILE.obs
								,years=years,period=period)
		
	theta.hat	<-	exp(optim.res$par[1])
	tau.hat		<-	exp(optim.res$par[2])
	b0.hat		<-	optim.res$par[3]
	b1.hat		<-	optim.res$par[4]

	
	neg.ll		<-	optim.res$value
	
	results		<-	c(tau=tau.hat,theta=theta.hat,b0=b0.hat,b1=b1.hat,loglik=-neg.ll)
	return(results)
}
```



```{r}
set.seed(123) 
b0_start	<-	runif(1,min=-36,max=-24) 
b1_start	<-	runif(1,min=0.656,max=0.984) 
theta_start	<-	log(runif(1,min=8,max=12)) 
tau_start	<-	log(runif(1,min=8,max=12))

SMILE_obs <- stochastic.sims
nreps <- length(SMILE_obs)

hat_mat			<-	array(0,dim=c(nreps,5),dimnames=list(1:nreps
															 ,c("tau","theta","b0","b1","loglik")))

for(i in 1:nreps){
	
	hat_mat[i,]	<-	SMILE.param.estim(b0 = b0_start,
	                                 b1 = b1_start,
	                                 theta = theta_start,
	                                 tau = tau_start,
	                                 SMILE.obs=SMILE_obs[[i]],period=3*52)
								}
```

```{r echo=FALSE}

# estimates from the stochastic simulations
hat_mat %>% as.data.frame() %>% rownames_to_column(var = "run") %>% 
  pivot_longer(-run, names_to = "parameter") -> long_hatmat

long_hatmat %>% 
  filter(parameter == "b0") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = b0, color = "red") +
  labs(title = "b0") +
  theme_classic() -> b0_plot

long_hatmat %>% 
  filter(parameter == "b1") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = b1, color = "red") +
  labs(title = "b1") +
  theme_classic() -> b1_plot

long_hatmat %>% 
  filter(parameter == "theta") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = theta, color = "red") +
  labs(title = "theta") +
  theme_classic() -> theta_plot

long_hatmat %>% 
  filter(parameter == "tau") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = tau, color = "red") +
  labs(title = "tau") +
  theme_classic() -> tau_plot

plot_grid(b0_plot, b1_plot, theta_plot, tau_plot, nrow = 2)

```

## with the new functions 

```{r eval=FALSE}
set.seed(123) 
b0_start	<-	runif(1,min=-36,max=-24) 
b1_start	<-	runif(1,min=0.656,max=0.984) 
theta_start	<-	log(runif(1,min=8,max=12)) 
tau_start	<-	log(runif(1,min=8,max=12))

SMILE_obs <- stochastic_sims
nreps <- length(SMILE_obs)

hat_mat			<-	array(0,dim=c(nreps,5),dimnames=list(1:nreps
															 ,c("tau","theta","b0","b1","loglik")))

for(i in 1:nreps){
	
	hat_mat[i,]	<-	SMILE_param_estim(b0 = b0_start,
	                                 b1 = b1_start,
	                                 theta = theta_start,
	                                 tau = tau_start,
	                                 SMILE.obs=SMILE_obs[[i]],period=3*52)
								}
```

Parameter estimates for each of the stochastic simulations we generated. All simulations used the values shown with the red line to be generated. 

```{r echo=FALSE}

# estimates from the stochastic simulations
hat_mat %>% as.data.frame() %>% rownames_to_column(var = "run") %>% 
  pivot_longer(-run, names_to = "parameter") -> long_hatmat

long_hatmat %>% 
  filter(parameter == "b0") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = b0, color = "red") +
  labs(title = "b0") +
  theme_classic() -> b0_plot

long_hatmat %>% 
  filter(parameter == "b1") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = b1, color = "red") +
  labs(title = "b1") +
  theme_classic() -> b1_plot

long_hatmat %>% 
  filter(parameter == "theta") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = theta, color = "red") +
  labs(title = "theta") +
  theme_classic() -> theta_plot

long_hatmat %>% 
  filter(parameter == "tau") %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  geom_vline(xintercept = tau, color = "red") +
  labs(title = "tau") +
  theme_classic() -> tau_plot

plot_grid(b0_plot, b1_plot, theta_plot, tau_plot, nrow = 2)

```
