---
title: "vax_manuscript"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
set.seed(9785)
library(SMILE)
library(tidyverse)
library(cowplot)
```

```{r}
theme_set(theme_bw())
```


```{r}
vax_seq <- seq(0, 1, 0.001)
my_beta_1 <- 5
my_beta_0 <- c(-1, 0, 2)

map(my_beta_0, \(.x) data.frame(surv_prob = calc_survival_prob(beta_0 = .x, beta_1 = my_beta_1, vax_rate = vax_seq),
                                beta_0 = .x, vax_rate = vax_seq)) |> 
  list_rbind() |> 
  ggplot(aes(x = vax_rate, y = surv_prob, color = factor(beta_0))) +
  geom_line() +
  scale_x_continuous(labels = scales::label_percent(), breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0,1,0.05)) +
  labs(y = "Survival probability", x = "Vaccination rate")
```


```{r}
# base parameters
my_years <- 10
my_tau <- 1
my_theta <- 100

# seasonal forcing parameters
my_b_fixed <- 0.001
my_b0 <- -30
my_b1 <- 0.85
my_outbreak_period <- 3*52

```


# Simulations 

```{r}
# no vaccine, three levels of survival
map(my_beta_0, \(.x) 
    smile_main(b0 = my_b0, b1 = my_b1, period = my_outbreak_period,
           theta = my_theta, tau = my_tau, years = my_years,
           vax = rep(0, my_years), beta_0 = .x, beta_1 = my_beta_1, output_df = TRUE) |> mutate(scenario = paste("beta_0 =", .x))) |> 
  list_rbind() -> sims_no_vax

head(sims_no_vax)
```


```{r}
sims_no_vax |> 
  pivot_longer(-c(week, scenario), names_to = "compartment", values_to = "n") |> 
  mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "Z"))) %>% 
    ggplot(aes(x = week, y = n, color = scenario)) +
    facet_wrap(~compartment, scales = "free", nrow = 2) +
    geom_line()
```

# Parameter space ----------------------------


```{r fig.width=12, fig.height=4, echo = FALSE, eval=FALSE}
# simplest scenario, no host dynamics
sim1 <- smile_main(tau = 10, theta = 100, b_fixed = 0.001, years = 20, age_struc = FALSE, N1 = 20000, zeta_novax = 0.5, output_df = TRUE)
sim1 %>% build_SMILE_plots(., 1) + labs(title = "Simplest - no host dynamics")
```
We start with a 20,000 susceptible population, but since there are no host dynamics, no reproduction, then the population just continues to decrease as a result of the outbreak. In this case, we have incorporated a high survival probability, exploring the range from 0.85 to 1, the lower the survival probability, the faster this decline in population size.

```{r fig.width=12, fig.height=4}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, b_fixed = 0.001, years = 20, age_struc = FALSE, N1 = 20000, zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "Z"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c()
```

Now we consider with population dynamics, where there is a yearly birth pulse, and we have death by other causes aside from just disease. Since we still have a relatively high survival probability, you will see that outbreaks occur yearly, following the birth pulses. Over time, the population tries to settle close to carrying capacity with yearly outbreaks.

```{r fig.width=13, fig.height=4}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, b_fixed = 0.001, years = 20, age_struc = TRUE, N1 = 20000, K =10000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
                  zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
   mutate(N = S + M) %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "N"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c() +
   labs(title = "K =10,000")
 
```

```{r fig.width=13, fig.height=4}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, b_fixed = 0.001, years = 20, age_struc = TRUE, N1 = 20000, K =20000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
                  zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
   mutate(N = S + M) %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "N"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c() +
   labs(title = "K =20,000")
 
```

```{r fig.width=13, fig.height=4}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, b_fixed = 0.001, years = 20, age_struc = TRUE, N1 = 20000, K =50000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
                  zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
   mutate(N = S + M) %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "N"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c() +
   labs(title = "K =50,000")
 
```

Now we incorporate seasonal transmission to the population. Given the 20 year time frame, we see that a K-=50,000 the population hasn't reached it's equilibrium yet, so we will stick to a K=20,000 to test the seasonal transmission. In  this case, population dynamics and survival don't play such a big role in outbreak timing, and we see that outbreaks are mainly dominated by the period in the seasonal transmission equation. Outbreaks are also larger, with the number of infecteds and deaths almost double what we saw in the scenarios with host dynamics and no seasonal forcing. The number of deaths is variable and changes according to the survival probability.

```{r fig.width=13, fig.height=4}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, years = 20, age_struc = TRUE, N1 = 20000, K =20000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
                  b0 = -30, b1 = 0.85, period = 3*52,
                  zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
   mutate(N = S + M) %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "N"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c() +
   labs(title = "b0 = -30, b1 = 0.85, period = 3*52")
 
```

We observe with this example how the period for seasonal transmission determines the outbreaks. It would be interesting to consider different types of outbreaks, perhaps thinking about dormant spores causing outbreaks 5+ years after the animal dies. 

```{r fig.width=13, fig.height=4}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, years = 20, age_struc = TRUE, N1 = 20000, K =20000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
                  b0 = -30, b1 = 0.85, period = 7*52,
                  zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
   mutate(N = S + M) %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "N"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c() +
   labs(title = "b0 = -30, b1 = 0.85, period = 7*52")
 
```

changing the parameters of the seasonal transmission function results in changes to the amplitude of the oscillations. The red line is the value used for no seasonality.

```{r fig.width= 8, fig.height=4}
nweeks <- 20*52
my_b_season <- b_season(b0 = -16, b1 = 0.6, period = 3*52,t= 0:nweeks)

data.frame(week = 0:nweeks, b = my_b_season) %>% 
  ggplot(aes(x = week, y = b)) +
  geom_path() +
  geom_hline(yintercept = my_b_fixed, color = "red") +
  theme_bw() +
  labs(y = "b - probability of transmission")
```

```{r fig.width=12, fig.height=10}
my_b0s <- c(-30, -25, -18)
my_b1s <- c(0.85, 0.7, 0.55)
period <- c(3,5, 7)
 
# we have 18 parameter combinations 

expand.grid(my_b0s, my_b1s, period) %>%  
  set_names(., c("b0", "b1", "period")) %>% 
  mutate(b = pmap(list(b0, b1, period), function(x,y,z) b_season(x,y,z*52, 0:nweeks))) %>%
  unnest(cols = c(b)) %>% 
  mutate(params = paste0("b0=", b0, ",b1=", b1, ",period=", period)) %>% 
  mutate(week = rep(0:nweeks, 27), 
         period = factor(period, levels = c(3,5, 7))) -> b_season_df

b_season_df %>% 
  # filter(b1 == 0.6) %>% 
  mutate(b0 = paste("b0 =", b0),
         b1 = paste("b1 =", b1)) %>% 
  ggplot(aes(x = week, y = b, group = params, color = period)) +
  facet_wrap(vars(b1, b0), ncol = 3, scales = "free") +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = my_b_fixed, color = "red") +
  theme_bw() +
  labs(y = "b - probability of transmission") +
  scale_color_manual(values = c("grey", "darkolivegreen", "slateblue"))

```


```{r fig.width=13, fig.height=4, echo=FALSE}
surv_prob_seq <- seq(0.85, 0.98, by = 0.01)

my_smile_df <- NULL
for(i in 1:length(surv_prob_seq)) {
  
  i_survival <- surv_prob_seq[i]
  sim <- smile_main(tau = 10, theta = 100, years = 20, age_struc = TRUE, N1 = 20000, K =20000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
                  b0 = -18, b1 = 0.55, period = 3*52,
                  zeta_novax = i_survival, output_df = TRUE) %>% 
    mutate(survival = i_survival)
  
  my_smile_df <- bind_rows(my_smile_df, sim)
}


 my_smile_df %>% 
   mutate(N = S + M) %>% 
    pivot_longer(-c(week, survival), names_to = "compartment", values_to = "n") %>% 
    mutate(compartment = factor(compartment, levels = c("S", "M", "I", "L", "E", "N"))) %>% 
    ggplot(aes(x = week, y = n, color = survival, group = survival)) +
    facet_wrap(~compartment, scales = "free", nrow = 1) +
    geom_line(linewidth = 1, alpha = 0.5) +
    theme_bw() +
   scale_color_viridis_c() +
   labs(title = "b0 = -18, b1 = 0.55, period = 3*52")
 
```


# IMGS2024


```{r}
# base parameters
my_years <- 23
my_tau <- 1
my_theta <- 100

# seasonal forcing parameters
my_b_fixed <- 0.001
my_b0 <- -30
my_b1 <- 0.85
my_outbreak_period <- 3*52

# Population dynamics
my_N1 <- 20000
my_K <- 20000
my_rho_pop <- 0.36
my_sigmaa <- 0.92^(1/52)


# my_params <- list(tau = 10, theta = 100, years = 20, age_struc = TRUE, N1 = 20000, K =20000, rho_pop = 0.36, sigmaa = 0.92^(1/52),
#                   b0 = -30, b1 = 0.85, period = 3*52,
#                   output_df = TRUE)

my_params <- list(tau = my_tau, theta = my_theta, years = my_years,
                  age_struc = TRUE, N1 = my_N1, K = my_K, rho_pop = my_rho_pop, sigmaa = my_sigmaa,
                  b0 = my_b0, b1 = my_b1, period = my_outbreak_period,
                  output_df = TRUE)

# Variable parameters
vax_seq <- seq(0, 1, 0.001)
my_beta_1 <- 5
my_beta_0 <- c(-1, 0, 2)
fixed_vax <- c(0, 0.1, 0.35, 0.5, 0.7)
random_vax <- runif(my_years,0, 0.8)
lao_cai_vax <- c(0.34, 0.24, 0.37, 0.44, 0.44, 0.64, 0.71, 0.69, 0.73, 
              0.72, 0.66, 0.60, 0.57, 0.27, 0.27, 0.16, 0.12, 0.12, 
              0.10, 0.08, 0.05, 0.05, 0.02)
```

```{r}
# function to visualize results fast
quick_facet <- function(params_list, my_title = NULL, return_df = FALSE) {
  
  sim_df <- do.call(smile_main, params_list)
  sim_b_season <- b_season(b0 = params_list$b0, b1 = params_list$b1, period = params_list$period, t = 0:(params_list$years*52))
  sim_R0 <- calc_local_R0(params_list$tau, params_list$theta,sim_b_season, sim_df$E)
  
  sim_df |> 
    bind_cols(b = sim_b_season, R0 = sim_R0) |> 
    mutate(Nt = S + M) |> 
    pivot_longer(-week, names_to = "compartment") |> 
    mutate(compartment = factor(compartment, levels = c("Nt", "S", "M", "I", "L", "E", "R0", "b", "Z"))) -> sim_df_complete
  
  if(return_df == TRUE) {
    return(sim_df_complete)
  } else {
    sim_df_complete |> 
      ggplot(aes(x = week, y = value)) +
      facet_wrap(~compartment, scales = "free") +
      geom_line() +
      labs(title =  my_title) +
      theme(legend.position = "none") +
      # scale_y_log10() +
      NULL -> my_smile_facet
    return(my_smile_facet)
  }
  
  
}

```


To start, assume a 50% survival rate for cattle, a coin toss if they are not vaccinated.

```{r}
novax_vax_vector <- rep(fixed_vax[1], my_years)
novax_params <- append(my_params, list(beta_1 = my_beta_1, beta_0 = my_beta_0[2], vax = novax_vax_vector))

quick_facet(novax_params, "No Vax - 50% survival")
```


Now, fixed vaccine at different levels, middle survival line.

```{r}

map(fixed_vax, function(.x){
  vax_vector <- rep(.x, my_years)
  vax_params <- append(my_params, list(beta_1 = my_beta_1, beta_0 = my_beta_0[2], vax = vax_vector))

  quick_facet(vax_params, return_df = TRUE) |> 
    mutate(fixed_vax_level = .x)
}) |> 
  list_rbind() -> fixed_vax_sims

fixed_vax_sims |> 
  mutate(fixed_vax_level = factor(fixed_vax_level)) |> 
  filter(compartment %in% c("S", "M", "I", "L", "E", "Z")) |> 
    ggplot(aes(x = week, y = value, group = fixed_vax_level, color = fixed_vax_level)) +
      facet_wrap(~compartment, scales = "free") +
      geom_line() +
      labs(title =  "Fixed vaccination")

```

```{r}
fixed_vax_sims |> 
  mutate(fixed_vax_level = factor(fixed_vax_level)) |> 
  filter(compartment %in% c("R0")) |> 
  ggplot(aes(x = week, y = value, color = fixed_vax_level, group = fixed_vax_level)) +
  facet_wrap(~fixed_vax_level, ncol=1) +
  geom_line(linewidth = 1.5, alpha = 0.8) +
  labs(title = "local R0") +
  geom_hline(yintercept = 1, color = "red", linetype = "dashed") +
  theme(legend.position = "none") -> R0_plot_fixed

fixed_vax_sims |> 
  mutate(fixed_vax_level = factor(fixed_vax_level)) |> 
  filter(compartment %in% c("L")) |> 
  ggplot(aes(x = week, y = value, color = fixed_vax_level, group = fixed_vax_level)) +
  facet_wrap(~fixed_vax_level, ncol=1) +
  geom_line(linewidth = 1.5, alpha = 0.8) +
  labs(title = "LIZ") +
  theme(legend.position = "none") -> LIZ_plot_fixed

plot_grid(R0_plot_fixed, LIZ_plot_fixed, ncol=2)


```



## Variable Vaccination

```{r}

data.frame(week = 1:(my_years*52), vax_rate = rep(random_vax, each=52)) |> 
  ggplot(aes(x = week, y = vax_rate)) +
  geom_line()

data.frame(year = 1:my_years, vax_rate = random_vax) |> 
  ggplot(aes(x = year, y = vax_rate)) +
  geom_line()
  
```


```{r}
map(my_beta_0, function(.x){
  
  vax_params <- append(my_params, list(beta_1 = my_beta_1, beta_0 = .x, vax = random_vax))

  quick_facet(vax_params, return_df = TRUE) |> 
    mutate(beta_0 = .x)
}) |> 
  list_rbind() -> random_vax_sims

random_vax_sims |> 
  filter(compartment %in% c("S", "M", "I", "L", "E", "Z")) |> 
    ggplot(aes(x = week, y = value, group = beta_0, color = factor(beta_0))) +
      facet_wrap(~compartment, scales = "free") +
      geom_line() +
      labs(title =  "Random vaccination")

```


```{r}
random_vax_sims |> 
  mutate(beta_0 = factor(beta_0)) |> 
  filter(compartment %in% c("R0")) |> 
  ggplot(aes(x = week, y = value, color = beta_0, group = beta_0)) +
  facet_wrap(~beta_0, ncol=1) +
  geom_line(linewidth = 1.5, alpha = 0.8) +
  labs(title = "local R0") +
  geom_hline(yintercept = 1, color = "red", linetype = "dashed") +
  theme(legend.position = "none") -> R0_plot_random

random_vax_sims |> 
  mutate(beta_0 = factor(beta_0)) |> 
  filter(compartment %in% c("L")) |> 
  ggplot(aes(x = week, y = value, color = beta_0, group = beta_0)) +
  facet_wrap(~beta_0, ncol=1, scales = "free_y") +
  geom_line(linewidth = 1.5, alpha = 0.8) +
  labs(title = "LIZ") +
  theme(legend.position = "none") -> LIZ_plot_random

random_vax_sims |> 
  mutate(beta_0 = factor(beta_0)) |> 
  filter(compartment %in% c("Z")) |> 
  ggplot(aes(x = week, y = value, color = beta_0, group = beta_0)) +
  facet_wrap(~beta_0, ncol=1) +
  geom_line(linewidth = 1.5, alpha = 0.8) +
  labs(title = "Z") +
  theme(legend.position = "none") -> Z_plot_random

plot_grid(R0_plot_random, LIZ_plot_random, Z_plot_random, ncol=3)
```
How does R0 change with survival due to vaccination?

```{r}
random_vax_sims |> 
  mutate(beta_0 = factor(beta_0)) |> 
  filter(compartment %in% c("R0", "Z")) |> 
  ggplot(aes(x = week, y = value, linetype = compartment, color = beta_0)) +
  facet_wrap(~beta_0, ncol = 1) +
  geom_line(linewidth = 1)
  
```

# Lao Cai

```{r}
data.frame(week = 1:(my_years*52), vax_rate = rep(lao_cai_vax, each=52)) |> 
  ggplot(aes(x = week, y = vax_rate)) +
  geom_line()
```


```{r}

map(my_beta_0, function(.x){
  
  vax_params <- append(my_params, list(beta_1 = my_beta_1, beta_0 = .x, vax = lao_cai_vax))

  quick_facet(vax_params, return_df = TRUE) |> 
    mutate(beta_0 = .x)
}) |> 
  list_rbind() -> lao_cai_vax_sims

lao_cai_vax_sims |> 
  # filter(compartment %in% c("S", "M", "I", "L", "E", "Z")) |> 
    ggplot(aes(x = week, y = value, group = beta_0, color = factor(beta_0))) +
      facet_wrap(~compartment, scales = "free") +
      geom_line() +
      labs(title =  "Lao Cai vaccination")

```

```{r}
lao_cai_vax_sims |> 
  mutate(beta_0 = factor(beta_0)) |> 
  filter(compartment %in% c("R0", "Z")) |> 
  ggplot(aes(x = week, y = value, linetype = compartment, color = beta_0)) +
  facet_wrap(~beta_0, ncol = 1) +
  geom_line(linewidth = 1) +
  labs(title = "R0 with Z") +
  geom_hline(yintercept = 1, color = "black", linetype = "dashed") -> A

lao_cai_vax_sims |> 
  mutate(beta_0 = factor(beta_0)) |> 
  filter(compartment %in% c("L")) |> 
  ggplot(aes(x = week, y = value, color = beta_0, group = beta_0)) +
  facet_wrap(~beta_0, ncol=1, scales = "free_y") +
  geom_line(linewidth = 1.5, alpha = 0.8) +
  labs(title = "LIZ") +
  theme(legend.position = "none") -> B

plot_grid(B, A, ncol = 2, rel_widths = c(1, 1.5), align = "h")
```

# Province-wide


```{r}
dien_bien <- c(0.300, 0.252, 0.256, 0.205, 0.241, 0.282, 0.269, 0.242, 0.178, 0.173)

# prov_params <- list(tau = my_tau, theta = my_theta,
#                   age_struc = FALSE, N1 = 100, sigmaa = 1,
#                   b0 = my_b0, b1 = my_b1, period = my_outbreak_period,
#                   output_df = TRUE)

prov_params <- list(tau = my_tau, theta = my_theta,
                  age_struc = TRUE, N1 = my_N1, K = my_K, rho_pop = my_rho_pop, sigmaa = my_sigmaa,
                  b0 = my_b0, b1 = my_b1, period = my_outbreak_period,
                  output_df = TRUE)


dien_bien_params <- append(prov_params, list(beta_1 = my_beta_1, beta_0 = my_beta_0[2], vax = dien_bien, years = length(dien_bien)))

do.call(smile_main, dien_bien_params)
quick_facet(dien_bien_params)

```
```{r}
quick_prov <- function(province_vax, plot_title){
  map(my_beta_0, function(.x){
  
  vax_params <- append(prov_params, list(beta_1 = my_beta_1, beta_0 = .x, vax = province_vax, years = length(province_vax)))

  quick_facet(vax_params, return_df = TRUE) |> 
    mutate(beta_0 = .x)
}) |> 
  list_rbind() |> 
    ggplot(aes(x = week, y = value, group = beta_0, color = factor(beta_0))) +
      facet_wrap(~compartment, scales = "free") +
      geom_line() +
      labs(title =  plot_title) -> A
  
  data.frame(week = 1:(length(province_vax)*52), vax_rate = rep(province_vax, each=52)) |> 
  ggplot(aes(x = week, y = vax_rate)) +
  geom_line() -> B
  
  plot_grid(A, B, ncol=1, rel_heights = c(3, 1))
}
```

```{r fig.width=10, fig.height=8}
dien_bien <- c(0.300, 0.252, 0.256, 0.205, 0.241, 0.282, 0.269, 0.242, 0.178, 0.173)
quick_prov(dien_bien, "Dien Bien")
```


```{r fig.width=10, fig.height=8}
ha_giang <- c(0.522, 0.410, 0.346, 0.337, 0.321, 0.550, 0.539, 0.538, 0.561, 0.575, 0.608, 0.692)
quick_prov(ha_giang, "Ha Giang")
```

```{r fig.width=10, fig.height=8}
lai_chau <- c(0.313, 0.271, 0.301, 0.124, 0.145, 0.200, 0.084, 0.143, 0.139, 0.138, 0.135, 0.131, 0.135, 0.139, 0.095, 0.187, 0.061, 0.046)
quick_prov(lai_chau, "Lai Chau")
```

```{r fig.width=10, fig.height=8}
son_la <- c(0.194, 0.171, 0.143, 0.110, 0.064, 0.039, 0.031, 0.023, 0.013, 0.004, 0.015, 0.026, 0.006, 0.007, 0.001, 0.000, 0.000, 0.000, 0.000)
quick_prov(son_la, "Son La")
```





```{r fig.width=10, fig.height=8}
lao_cai <- c(0.000, 0.130, 0.093, 0.144, 0.174, 0.173, 0.259, 0.290, 0.284, 0.305, 0.303, 0.000, 0.263, 0.253, 0.121, 0.123, 0.112, 0.055, 0.055,
             0.045, 0.035, 0.023, 0.021, 0.011, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000, 0.000)
quick_prov(lao_cai, "Lao Cai")
```

```{r fig.width=10, fig.height=8}
cao_bang <- c(0.052, 0.056, 0.051, 0.042, 0.040, 0.024, 0.023)
quick_prov(cao_bang, "Cao Bang")
```




